Reanalysis Step 2: Multiple Imputation

### 2025-08-30 reanalysis-2-multiple_imputation.qmd
simplify

# Convergence plot
pdf(file.path(reanalysis_data_dir, "mice_convergence.pdf"))
plot(mice_results)
dev.off()
```

## Imputation Diagnostics (selected variables)

```{r imputation-diagnostics}
# Show diagnostic plots only when a variable has missingness
if (sum(is.na(df_mice$CCI)) > 0) {
  densityplot(mice_results, ~ CCI, main = "CCI") |> print()
  stripplot(mice_results, CCI ~ .imp, pch = 20, cex = 1.0, xlab = "Dataset (1=Original, 2+=Imputed)", main = "CCI: Original vs Imputed") |> print()
}
if (sum(is.na(df_mice$DCSI)) > 0) {
  densityplot(mice_results, ~ DCSI, main = "DCSI") |> print()
  stripplot(mice_results, DCSI ~ .imp, pch = 20, cex = 1.0, xlab = "Dataset (1=Original, 2+=Imputed)", main = "DCSI: Original vs Imputed") |> print()
}
if (sum(is.na(df_mice$baseline_VA_logMAR)) > 0) {
  densityplot(mice_results, ~ baseline_VA_logMAR, main = "Baseline VA (logMAR)") |> print()
  stripplot(mice_results, baseline_VA_logMAR ~ .imp, pch = 20, cex = 1.0, xlab = "Dataset (1=Original, 2+=Imputed)", main = "Baseline VA: Original vs Imputed") |> print()
}
if (sum(is.na(df_mice$intermediate_VA_logMAR)) > 0) {
  densityplot(mice_results, ~ intermediate_VA_logMAR, main = "Intermediate VA (logMAR)") |> print()
  stripplot(mice_results, intermediate_VA_logMAR ~ .imp, pch = 20, cex = 1.0, xlab = "Dataset (1=Original, 2+=Imputed)", main = "Intermediate VA: Original vs Imputed") |> print()
}
if (sum(is.na(df_mice$outcome_VA_logMAR)) > 0) {
  densityplot(mice_results, ~ outcome_VA_logMAR, main = "Outcome VA at 2 Years (logMAR)") |> print()
  stripplot(mice_results, outcome_VA_logMAR ~ .imp, pch = 20, cex = 1.0, xlab = "Dataset (1=Original, 2+=Imputed)", main = "Outcome VA: Original vs Imputed") |> print()
}

# person_dr (ordinal categorical): summarize proportions across imputations
if (sum(is.na(df_mice$person_dr)) > 0) {
  dr_props <- sapply(1:n_imputations, function(i) {
    imp_data <- complete(mice_results, i)
    prop.table(table(imp_data$person_dr))
  })
  if (!is.null(dim(dr_props))) {
    dr_summary <- data.frame(
      Category = rownames(dr_props),
      Mean_Proportion = apply(dr_props, 1, mean) |> round(3),
      SD = apply(dr_props, 1, sd) |> round(3)
    )
    kable(dr_summary, caption = "DR severity proportions across imputations")
  }
}

# Categorical example: cataract surgery proportions across imputations
if (sum(is.na(df_mice$catsurg_before_hitplus_cat)) > 0) {
  cat_props <- sapply(1:n_imputations, function(i) {
    imp_data <- complete(mice_results, i)
    prop.table(table(imp_data$catsurg_before_hitplus_cat))
  })
  if (!is.null(dim(cat_props))) {
    cat_summary <- data.frame(
      Level = rownames(cat_props),
      Mean_Proportion = apply(cat_props, 1, mean) |> round(3),
      SD = apply(cat_props, 1, sd) |> round(3)
    )
    kable(cat_summary, caption = "Cataract surgery proportions across imputations")
  }
}
```

>>>>

# Minimal completeness verification: first completed dataset has no missing values
first_imp <- complete(mice_results, 1)
missing_after <- colSums(is.na(first_imp))
missing_after <- missing_after[missing_after > 0]
if (length(missing_after) > 0) {
  log_message("WARNING: Missing values remain after imputation:")
  print(missing_after)
} else {
  log_message("SUCCESS: All missing data imputed in first completed dataset.")
}
```
